---
title: "Multiclass Regression"
author: "Joshua Chen"
date: "12/2/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Loading packages

```{r, result = 'hide'}
library(bayestestR)
library(ggplot2)
library(mombf)
library(tidyverse)
library(readr)
library(glmnet)
library(caret)
```
```{r}
softmax <- function(x) {
    exp_x <- exp(x - max(x))  # Subtracting max(x) for numerical stability
    return(exp_x / sum(exp_x))
}
```


```{r, results = 'hide'}
X_train <- read_csv("processed_data/X_test.csv")
X_test <- as.matrix(read_csv("processed_data/X_train.csv"))
y_train_dum <- read_csv("processed_data/y_test_dum.csv")
y_test_dum <- read_csv("processed_data/y_train_dum.csv")
```

```{r}
scaling_model <- preProcess(X_train, method = c("center", "scale"))
X_train <- predict(scaling_model, newdata = X_train)
X_test <- predict(scaling_model, newdata = X_test)
```

## Simple Logistic Regression
```{r}
set.seed(31415)
class = colnames(y_test_dum)
lr_result = list()

for (c in class) {
  # Subset the target variable for the current class
  y_train_class <- y_train_dum[[c]]
  
  # Fit LASSO logistic regression model using cross-validation
  model_name <- paste('lr_', c, sep = "")
  X_train_matrix <- as.matrix(X_train)
  assign(model_name, glm(y_train_class ~ ., data = data.frame(cbind(y_train_class, X_train_matrix)), family = "binomial"))
  
  # Predict on the testing set
  fit.lr <- get(model_name)
  pred <- predict(fit.lr, newdata = as.data.frame(X_test),  type = "response")
  
  # Store the predictions in the result list
  lr_result[[c]] <- pred
}
```
```{r}
lr_df <- as.data.frame(lr_result)
colnames(lr_df) = class
head(lr_df)
```


```{r}
lr_df_binary <- apply(lr_df, 1, function(row) {
  binary_row <- as.numeric(row == max(row))
  names(binary_row) <- colnames(lr_df)
  return(binary_row)
})

lr_df_binary <- as.data.frame(t(lr_df_binary))
head(lr_df_binary)
```


```{r}
max_pred_names <- as.factor(colnames(lr_df_binary)[max.col(lr_df_binary, 'first')])
max_test_names <- as.factor(colnames(y_test_dum)[max.col(y_test_dum, 'first')])
conf_matrix = confusionMatrix(max_pred_names, max_test_names)
conf_matrix
```

## 10 Fold LASSO CV

```{r}
set.seed(31415)
class = colnames(y_test_dum)
lasso_result = list()

for (c in class) {
  # Subset the target variable for the current class
  y_train_class <- y_train_dum[[c]]
  
  # Fit LASSO logistic regression model using cross-validation
  model_name <- paste('lasso_', c, sep = "")
  assign(model_name, cv.glmnet(x = as.matrix(X_train), y = y_train_class, family = "binomial", alpha = 1, nfolds = 20))
  
  # Display the optimal lambda value chosen by cross-validation
  fit.lasso <- get(model_name)
  best_lambda <- fit.lasso$lambda.min
  cat("Optimal Lambda for class", c, ":", best_lambda, "\n")
  
  # Predict on the testing set
  predictions <- predict(fit.lasso, newx = as.matrix(X_test), s = best_lambda, type = "response")
  
  # Store the predictions in the result list
  lasso_result[[c]] <- predictions
}
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
lasso_df <- as.data.frame(lasso_result)
colnames(lasso_df) = class
head(lasso_df)
```

Selecting Maximum Probability as 1 and 0s elsewhere
```{r}
lasso_df_binary <- apply(lasso_df, 1, function(row) {
  binary_row <- as.numeric(row == max(row))
  names(binary_row) <- colnames(lasso_df)
  return(binary_row)
})

lasso_df_binary <- as.data.frame(t(lasso_df_binary))
head(lasso_df_binary)
```


```{r}
lassomax_pred_names <- as.factor(colnames(lasso_df_binary)[max.col(lasso_df_binary, 'first')])
lassomax_test_names <- as.factor(colnames(y_test_dum)[max.col(y_test_dum, 'first')])
lassoconf_matrix = confusionMatrix(lassomax_pred_names, lassomax_test_names)
lassoconf_matrix
```

Summary Statistics:
```{r}
accuracy <- conf_matrix$byClass[, "Balanced Accuracy"]
precision <- conf_matrix$byClass[, "Pos Pred Value"]  # Precision
recall <- conf_matrix$byClass[, "Sensitivity"]        # Recall
f1_score <- conf_matrix$byClass[, "F1"]
accuracy
```

```{r}
precision
```

## LASSO-BIC
```{r}
lasso.bic.logistic <- function(y,x,extended=FALSE) {
  #Select model in LASSO path with best BIC (using LASSO regression estimates)
  #Input
  # - y: vector with response variable
  # - x: design matrix
  #
  #Output: list with the following elements
  # - coef: LASSO-estimated regression coefficient with lambda set via BIC
  # - ypred: predicted y
  # - lambda.opt: optimal value of lambda
  # - lambda: data.frame with bic and number of selected variables for each value of lambda
  require(glmnet)
  fit <- glmnet(x=x,y=y,family='binomial',alpha=1)
  pred <- predict(fit,newx=x,type='response')
  n <- length(y)
  p <- colSums(fit$beta!=0) + 1
  if (!extended){
    bic <- -2* colSums(y*log(pred)+(1-y)*log(1-pred)) + log(n)*p 
  } else {
    bic <- -2* colSums(y*log(pred)+(1-y)*log(1-pred)) + log(n)*p + 2*log(choose(ncol(x),p))
  }
  sel <- which.min(bic)
  beta <- c(fit$a0[sel],fit$beta[,sel]); names(beta)[1]= 'Intercept'
  ypred <- pred[,sel]
  ans <- list(model=fit,coef=beta,ypred=ypred,lambda.opt=fit$lambda[sel],lambda=data.frame(lambda=fit$lambda,bic=bic,nvars=p))
  return(ans)
}
```

```{r}
set.seed(31415)
class = colnames(y_test_dum)
bic_result = list()

for (c in class) {
  # Subset the target variable for the current class
  y_train_class <- y_train_dum[[c]]
  
  # Fit LASSO logistic regression model using cross-validation
  model_name <- paste('lassobic_', c, sep = "")
  assign(model_name, lasso.bic.logistic(y=y_train_class,x=as.matrix(X_train)))
  
  # Display the optimal lambda value chosen by cross-validation
  fit.lassobic <- get(model_name)
  best_lambda <- fit.lassobic$lambda.opt
  cat("Optimal Lambda for class", c, ":", best_lambda, "\n")
  
  # Predict on the testing set
  predictions <- predict(fit.lassobic$model, newx = X_test, s = best_lambda, type = 'response')
  
  # Store the predictions in the result list
  bic_result[[c]] <- predictions
}
```

```{r}
bic_df <- as.data.frame(bic_result)
colnames(bic_df) = class
head(bic_df)
```

Selecting Maximum Probability as 1 and 0s elsewhere
```{r}
bic_df_binary <- apply(bic_df, 1, function(row) {
  binary_row <- as.numeric(row == max(row))
  names(binary_row) <- colnames(bic_df)
  return(binary_row)
})

bic_df_binary <- as.data.frame(t(bic_df_binary))
head(bic_df_binary)
```


```{r}
bicmax_pred_names <- as.factor(colnames(bic_df_binary)[max.col(bic_df_binary, 'first')])
bicmax_test_names <- as.factor(colnames(y_test_dum)[max.col(y_test_dum, 'first')])
bicconf_matrix = confusionMatrix(bicmax_pred_names, bicmax_test_names)
bicconf_matrix
```

## Bayesian Model Selection
```{r}
bayes_results = list()

for (c in class) {
  # Subset the target variable for the current class
  y_train_class <- y_train_dum[[c]]
  
  # Fit Bayes model with Zellnor's non-information prior
  model_name <- paste('bayes_', c, sep = "")
  assign(model_name, modelSelection(y_train_class ~ ., data=X_train, priorCoef = zellnerprior(taustd =  1),family = 'binomial', priorDelta=modelbbprior(1,1)))
  print(paste(c,":", " model fitted", sep = ""))
  # Display the optimal lambda value chosen by cross-validation
  fit.bayes <- get(model_name)

  
  # Predict on the testing set
  predictions <- predict(fit.bayes, newdata = X_test,data = X_train, type= 'response')
  
  # Store the predictions in the result list
  bayes_results[[c]] <- predictions
}
```


```{r}
bayes_means <- lapply(bayes_results, function(df) df[, 1])
bayes_results_df <- data.frame(bayes_means)
head(bayes_results_df)
```


```{r}
bayes_df_binary <- apply(bayes_results_df, 1, function(row) {
  binary_row <- as.numeric(row == max(row))
  names(binary_row) <- colnames(bayes_results_df)
  return(binary_row)
})

bayes_df_binary <- as.data.frame(t(bayes_df_binary))
head(bayes_df_binary)
```

Confusion Matrix：
```{r}
bayesmax_pred_names <- as.factor(colnames(lasso_df_binary)[max.col(bayes_df_binary, 'first')])
bayesmax_test_names <- as.factor(colnames(y_test_dum)[max.col(y_test_dum, 'first')])
bayes_conf_matrix = confusionMatrix(bayesmax_pred_names, bayesmax_test_names)
bayes_conf_matrix
```

